---
title: "cheatsheet"
author: "emily"
date: "2023-01-31"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### First Things First!
```{r}
#library(tidyverse)
#install.packages("tidyverse")
#library(janitor)
#install.packages("janitor")
```

## Formating Funsies
#title
_italics_ 
**bold**
insert website links:
[webpage name](website URL)
insert email links:
[person name](mailto:emailaddress)

###Lab One 
##get your working directory!
```{r}
getwd()
```
##basic arithmetic
```{r}
4+4
4*4
(4+2)/3
4+3/3
```
##Assigning vectors and how to use 'em
```{r}
x <- c(1, 2, 3, 4, 5, 6) #c stands for concatenate, but basically means combine
```
```{r}
mean(x)
median(x)
sd(x)
```

### Lab Two
##objects and what to do with 'em
```{r}
x <- 5
#read as: x gets 5
#note: unlike vector, there is no c or multitude of #s, just one
```

```{r}
objects_can_also_be_words <- 4
```
```{r}
sum (x, objects_can_also_be_words)
```
```{r}
add_objects <- x + objects_can_also_be_words
```

```{r}
add_objects_diff <- sum (x, objects_can_also_be_words)
```

## Types of Data
```{r}
my_numeric <- 42
my_integer <- 2L #adding an L automatically denotes an integer
my_character <- "universe"
my_logical <- FALSE
my_complex <- 2+4i
```
#What kind of data am I working with?
```{r}
class(my_numeric) #or you can do:
is.integer(my_numeric) #is my_numeric an integer?
```

## NAs, Missing Data
```{r}
my_missing <- NA
```
```{r}
is.na(my_missing)
```
```{r}
anyNA(my_missing)
```
```{r}
mean(new_vector, na.rm=T) #na.rm=T removes the NA values in the vector
```

## More Vector Stuff
```{r}
my_vector <- c(10, 20, 30) #numeric vector
```
```{r}
days_of_the_week <- c("Monday", "Tuesday", "Wednesday", "Thrusday", "Friday", "Saturday", "Sunday") #charactor vector, because all characters have quotes and can be called "strings"
```
```{r}
my_vector_sequence <- c(1:100) #assigns vector the numbers 1 to 100
```
#Vector Elements
```{r}
days_of_the_week[3] #[] to pull out elements in the vector as long as their position is specified (3 = third input = wednesday)
```

## Data Matrices
```{r}
Philosophers_Stone <- c(317.5, 657.1) #data matrices is a series of stacked vectors
Chamber_of_Secrets <- c(261.9, 616.9)
Prisoner_of_Azkaban <- c(249.5, 547.1)
Goblet_of_Fire <- c(290.0, 606.8)
Order_of_the_Phoenix <- c(292.0, 647.8)
Half_Blood_Prince <- c(301.9, 632.4)
Deathly_Hallows_1 <- c(295.9, 664.3)
Deathly_Hallows_2 <- c(381.0, 960.5)
```
```{r}
box_office <- c(Philosophers_Stone, Chamber_of_Secrets, Prisoner_of_Azkaban, Goblet_of_Fire, Order_of_the_Phoenix, Half_Blood_Prince, Deathly_Hallows_1, Deathly_Hallows_2) 
#combine all your vectors into a new object
```
```{r}
harry_potter_matrix <- matrix(box_office, nrow = 8, byrow = T)
#use <- matrix function to assign your object to create a new matrix, and tell it how to organize with nrow and byrow
```
```{r}
region <- c("US", "non-US") #create new vectors to use for naming
titles <- c("Philosophers_Stone", "Chamber_of_Secrets", "Prisoner_of_Azkaban", "Goblet_of_Fire", "Order_of_the_Phoenix", "Half_Blood_Prince", "Deathly_Hallows_1", "Deathly_Hallows_2")
```
```{r}
colnames(harry_potter_matrix) <- region #name your column and row with your vectors
rownames(harry_potter_matrix) <- titles
```
```{r}
#print that bitch out!
harry_potter_matrix
```

#Using a Data Matrix
```{r}
global <- rowSums(harry_potter_matrix) #you can calculate the total earnings of each movie
```
```{r}
harry_potter_matrix[2,1] #we can select values or a series of values with a data matrix
harry_potter_matrix[1:4] #we can select a series a values with :
non_us_earnings <- all_harry_potter_matrix[ ,2]
mean(non_us_earnings) #select whole columns and do math!
```

#Adding Rows and Columns
```{r}
all_harry_potter_matrix <- cbind(harry_potter_matrix, global)
#note we are making a new object to reflect the addition of a new column with the cbind
```
```{r}
total_earnings <- colSums(all_harry_potter_matrix)
final_harry_potter_matrix <- rbind(all_harry_potter_matrix, total_earnings) 
#create the new object, then add a new row via rbind
```

### Lab Three
##Data Frame
```{r}
Sex <- c("male", "female", "male") #let's make some vectors to put in the data frame
Length <- c(3.2, 3.7, 3.4)
Weight <- c(2.9, 4.0, 3.1)
```
```{r}
hbirds <- data.frame(Sex, Length, Weight) #BOOM data frame via <- data.frame
```

**note** Data Matrix vs Data Frame?
Data Matrix is cleaner, identifies the class of variables, info about the class
dbl: double precision floating point

#analyzing a data frame
```{r}
names(hbirds) #these do the same thing
colnames(hbirds)
```
```{r}
dim(hbirds) #dimensions and structure
str(hbirds)
```
```{r}
hbirds <- data.frame(sex = Sex, length_in = Length, weight_oz = Weight) #renaming
```
```{r}
w <- hbirds$weight_oz # use $ to select values in a whole column
mean(w)
```
```{r}
hbirds$neighborhood <- c("lakewood", "brentwood", "lakewood", "scenic Heights")
hbirds
#add rows like with a data matrix: make new vector and rbind
#with columns, we can use the $
```


## .csv Files
```{r}
write.csv(hbirds, "what_you_want_for_the_name.csv", row.names = FALSE)
# use this to save as .csv file
# row.names = FALSE means row numbers won't be printed
```
```{r}
readr::read_csv("hsprings_data.csv")
# to import the data; you can control where Rstudio looks by adding "data/" to tell it to look in the data folder.
hot_springs <- readr::read_csv("hsprings_data.csv")
#you can aslo assign it to an object
```
#Analyzing .csv Files
```{r}
class(hot_springs$scientist) # what class is this column of data?
hot_springs$scientist <- as.factor(hot_springs$scientist) #changing the class to factor
levels(hot_springs$scientist) #showing the levels of the factor
```
```{r}
summary(filename)
```
```{r}
glimpse(filename)
```
```{r}
nrow(filename) #the number of rows or observations
```
```{r}
ncol(filename) #the number of columns or variables
```
```{r}
head(filename, n = 10) #prints first n rows of data
```
```{r}
tail(filename, n = 10) #prints last n rows of data
```
```{r}
table(.csv_object$variable) #makes a table using that variable in the file, produces counts
```
```{r}
View(fish) #opens the table, or you can do so in environment tab
```

### Lab Four
# Filter and Select
```{r}
object <- filter(filename, observation) #pulls out observations (rows) that meet specific criteria
```
```{r}
select(filename, "column1", "column2") #select pulls out columns of interest
```
```{r}
select(filename, start_col:end_col) #select a range of columns
```
```{r}
select(filename, -colno, -colbad, -colwrong) # - selects everything except specified
```
```{r}
select(filename, contains("word")) #only select variables with "word" in it
```
```{r}
select(filename, starts_with("word")) #take a wild guess.
```
```{r}
select(filename, matches("a.+er")) #select variable containing letter + subsequent string like "er"
```
```{r}
select_if(fish, is.numeric) #select based on class of data
```
```{r}
select_if(fish, ~!is.numeric(.)) #select all columns NOT a class of data

```
```{r}
filter(object, column == "observation") #can use & for multiple things to pull, or | for or
```
```{r}
filter(object, column %in% c(rowvalue)) #Filtering multiple values within the same variable requires the `%in%`
```
```{r}
filter(object, between(column, rowvalue, rowvalue)) #use `between` if you are looking for a range of specific values
```
```{r}
filter(object, near(column, rowvalue, tol = your_tolerance)) #note: rowvalue has to be a numeric

```

**note**
filter = rows = observations
select = columns = variables
filter operators: >, >=, <, <=, != (not equal), and == (equal)
! excludes specific observations
1. ends_with() = Select columns that end with a character string  
2. contains() = Select columns that contain a character string  
3. matches() = Select columns that match a regular expression  
4. one_of() = Select columns names that are from a group of names 

#Organize Within the Data Frame
```{r}
object <- select_all(data.frame, tolower)
#makes column names all lower case; also works with toupper
```
```{r}
select_all(object, ~str_replace(., " ", "_"))
#replaces funky shit like blank spaces
```

### Lab Five
##Pipes via ctrl + shift + m
Feeds the output from one function into the input of another function. This helps us keep our code sequential and clean
```{r}
mammals %>% #example from lab 5
  select(family, genus, species, gestation) %>% 
  filter(gestation>=15)
```
## Arrange
```{r}
fish %>%  #example from lab 5
  select(lakeid, scalelength) %>% 
  arrange(desc(scalelength)) # desc means large -> small, but usually its from small -> large
```
## Mutate
allows us to create a new column from existing columns in a data frame
```{r}
fish %>% #used mutate to create a new length column out of an old one
  mutate(length_mm = length*10) %>% 
  select(fish_id, length, length_mm)
```
# Mutate_all
```{r}
mammals %>%
  mutate_all(tolower)
```
useful for cleaning data
```{r}
mammals %>% 
  mutate(across(c("order", "family"), tolower))
```
Using the across function we can specify individual columns.
# Ifelse
```{r}
mammals %>% 
  select(genus, species, newborn) %>%
  mutate(newborn_new = ifelse(newborn == -999.00, NA, newborn))%>% 
  arrange(newborn)
```
you first specify a logical statement, afterwards what needs to happen if the statement returns `TRUE`, and lastly what needs to happen if it's  `FALSE`

##Janitor
```{r}
  janitor::clean_names(variable)
```
##Tabyl
```{r}
tabyl(object, column) #produces counts but also percentages
```
## Big Fat Juicy Example
Let's do a bit more practice to make sure that we understand `select()`, `filter()`, and `mutate()`. Start by building a new data frame `msleep24` from the `msleep` data that: contains the `name` and `vore` variables along with a new column called `sleep_total_24` which is the amount of time a species sleeps expressed as a proportion of a 24-hour day. Remove any rows with NA's and restrict the `sleep_total_24` values to less than 0.3. Arrange the output in descending order.  
```{r}
msleep24 <- msleep %>% 
  mutate(sleep_total_24 = sleep_total/24) %>% 
  select(name, vore, sleep_total_24) %>% 
  filter(!is.na(vore)) %>% #removing NAs from a variable
  filter(sleep_total_24<=.3) %>% 
  arrange(desc(sleep_total_24))
```

### Lab Six
```{r}
summarise() #literally what do you think it does bro
```
```{r}
skim() #find # of unique observations within a variable
```
```{r}
hist(object$observation) #makes a histogram of ur stuff
```
#summarize example
```{r}
msleep %>% 
  filter(order == "Primates") %>% 
  summarize(mean_bodywt = mean(bodywt), 
              min_bodywt = min(bodywt),
              max_bodywt = max(bodywt),
            total = n())
```

