---
title: "Lab 6 Intro"
date: "`r Sys.Date()`"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library("jpeg")
```

## Setup
1. Login to the lab computer (please don't use your personal computer).  
2. Navigate to github.com and login.   
2. Use GitHub desktop to clone your repository to the desktop.   
5. Copy the class repository to the desktop (https://github.com/jmledford3115/datascibiol).  
6. Copy the files for today's lab from the class repository and paste them into **your** repository.  
7. Open today's lab in RStudio.  

## Review from last time
### *With a partner, discuss the following questions*
1. What is a pipe? Why are they useful in R?
A pip is a feature that allows you to feed the output from one function into the input of another function. This is usefel in R because it keeps your code clean and efficient.

2. What is the shortcut for making a pipe?
ctrl + shift + m

3. What is the difference between `select`, `filter`, and `mutate`?
'Select' operates using columns, 'filter' via rows, and 'mutate' allows us to make new columns in the dataframe. When combined with 'ifelse', 'mutate' can also replace values in a dataframe. 

4. What are the logistics of midterm 1?
first 50 mins: work by yourself, then push to github
next 50 mins: work with others, then push to github

### Warm-up
1. Open the data `ecs21351-sup-0003-SupplementS1.csv`
```{r}
library(tidyverse)
library(janitor)
```
```{r}
ecosphere <- readr::read_csv("data/ecs21351-sup-0003-SupplementS1.csv", skip =2)
```

2. Clean the names of the variables
```{r}
colnames(ecosphere)
```
```{r}
ecosphere <- janitor::clean_names(ecosphere)
ecosphere
```

3. Explore the data using a function of your choice
```{r}
glimpse(ecosphere)
```

4. Are there more veggie, omni, insecivore, or nectar feeders in the data?
```{r}
ecosphere %>% 
  tabyl(diet) %>% 
  arrange(desc(n))
```

5. Which families are Veggie?
```{r}
ecosphere %>% 
  filter(diet == "Vegetation") %>% 
  select(family, diet) %>% 
  tabyl(diet)
```

